### 2015-11-16 cheatsheets/ode_formal_slns.tex
Added ODE stuff

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{array}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\title{Formal Solutions To ODE's}
\date{}
\begin{document}
\maketitle

\section{Homogeneous Solutions}

\subsection{First Order}

We have the following system
\[
	\begin{cases}
	y'+ay=0\\
	y(0)=\alpha
	\end{cases}
\]

\end{document}