### 2015-11-16 cheatsheets/ode_formal_slns.tex
Added homogeneous solutions

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{array}
\usepackage{hyperref}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\title{Formal Solutions To ODE's}
\date{}
\begin{document}
\maketitle

\section{Homogeneous Solutions}

\subsection{First Order}

We have the following differential equation
\[
	y'+ay=0.
\]
Then
\[
	y=c e^{-at},
\]
where $c$ is determined by the initial condition.

\subsection{Second Order}

We have the following differential equation
\begin{equation}
	y''+ay'+by=0. \label{eqn1}
\end{equation}
Then we solve the following equation for $r$,
\[
	r^2+ar+b=0.
\]

If $r=\alpha\pm i\beta$, then the general solution is
\[
	y=c_1 e^{\alpha x}\cos(\beta x)+c_2 e^{\alpha x}\sin(\beta x)
\]
where $c_1$ and $c_2$ are determined by the initial conditions.

If \autoref{eqn1} has two real solutions, $r_1$ and $r_2$, then
\[
	y=c_1e^{r_1 x}+c_2 e^{r_2x}
\]
where $c_1$ and $c_2$ are determined by the initial conditions.

If \autoref{eqn1} has a real repeated root, $r$, then
\[
	y=c_1e^{rx}+c_2xe^{rx}
\]
where $c_1$ and $c_2$ are determined by the initial conditions.

\end{document}