### 2015-11-16 cheatsheets/ode_formal_slns.tex
Added nonhomogeneous solutions

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{array}
\usepackage{hyperref}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\title{Formal Solutions To ODE's}
\date{}
\begin{document}
\maketitle

\section{Homogeneous Solutions}

\subsection{First Order}

We have the following differential equation
\[
	y'+ay=0.
\]
Then
\[
	y=c e^{-at},
\]
where $c$ is determined by the initial condition.

\subsection{Second Order}

We have the following differential equation
\begin{equation}
	y''+ay'+by=0. \label{eqn1}
\end{equation}
Then we solve the following equation for $r$,
\[
	r^2+ar+b=0.
\]

If $r=\alpha\pm i\beta$, then the general solution is
\[
	y=c_1 e^{\alpha x}\cos(\beta x)+c_2 e^{\alpha x}\sin(\beta x)
\]
where $c_1$ and $c_2$ are determined by the initial conditions.

If \autoref{eqn1} has two real solutions, $r_1$ and $r_2$, then
\[
	y=c_1e^{r_1 x}+c_2 e^{r_2x}
\]
where $c_1$ and $c_2$ are determined by the initial conditions.

If \autoref{eqn1} has a real repeated root, $r$, then
\[
	y=c_1e^{rx}+c_2xe^{rx}
\]
where $c_1$ and $c_2$ are determined by the initial conditions.

\section{Nonhomogeneous}

Suppose we have the following differential equation.
\begin{equation}\label{eqn2}
	ay'' + by' +cy = f(x)
\end{equation}

We begin by finding the solution to the homogeneous equation,
\[
	ay'' + by' +cy = 0.
\]
Call this solution the general solution or $y_g$.
Note, do \textbf{NOT} apply any initial conditions for this solution.

We now have a table for $f(x)$ and the form of the particular solution.
\[
\begin{array}{|c|c|}
\hline
f & y_p\\\hline
x & Ax+B\\\hline
x^2 & Ax+B+C\\\hline
e^x & Ae^x\\\hline
\cos(n\pi x) \text{ or } \sin(n\pi x) & A\cos(n\pi x) + B\sin(n\pi x)\\\hline
x\cos(n\pi x) \text{ or } \sin(n\pi x) & (Ax+B)\cos(n\pi x) + (Cx +D)\sin(n\pi x)\\\hline
\end{array}
\]
To find the $A,B,C,D$, plug $y_p$ back into \autoref{eqn2}, and match terms with $f(x)$.
The solution to \autoref{eqn2} is $y=y_g+y_p$.
Now apply your initial conditions to determine your as of now undetermined coefficients.

\end{document}