### 2015-11-21 prettyproofs/proofs.tex
presentation

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{array}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\NN}{\mathbb{N}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\CC}{\mathbb{C}}

\newcommand{\Proof}{\textit{\textbf{Proof: }}}
\newcommand{\scc}{\begin{cases*}}
\newcommand{\ecc}{\end{cases*}}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{axiom}{Axiom}
\newtheorem{proof}{Proof}

\title{Proofs}
\date{}
\begin{document}
\maketitle
\hrule
\noindent\\\\
\begin{axiom}
The number $0.99999... = 1.00000...$.
\end{axiom}
\begin{theorem}
There is no bijection between the natural numbers $\NN$ and the set $(0,1)$.
\end{theorem}
\noindent
\indent \Proof Suppose $|\NN| = |(0,1)|$. 
Let $f : \NN \to (0,1)$ be a bijection. 
We will let $a_{i} = f(i)$ and $a_{ij}$ be the $j^{\text{th}}$ decimal value of $a_{i}$. 
We can construct the bijection by
\begin{alignat*}{3}
a_{1} &= .a_{11}a_{12}a_{13}a_{14}\dots\ &&= f(a_{1})\\
a_{2} &= .a_{21}a_{22}a_{23}a_{24}\dots &&= f(a_{2})\\
a_{3} &= .a_{31}a_{32}a_{33}a_{34}\dots &&= f(a_{3})\\
a_{4} &= .a_{41}a_{42}a_{43}a_{44}\dots &&= f(a_{4})\\
\dots
\end{alignat*}
Now we can take an element $b \in (0,1)$ such that $b_{i}$ is the $i^{\text{th}}$ digit of $b$. 
We want to construct the digits of $b$ such that each space along the $a_{ii}$ diagonal is not equal to a digit in $b$. We can express this by
\[
b_{i} = 
\scc
n,\qquad a_{i} \neq n\\
n+1,\ a_{i} = n\\
8,\qquad\ a_{i} = 9
\ecc
\]
\end{document}