### 2015-12-03 comporg/review.tex

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage[table]{xcolor}
\usepackage{array}
\usepackage{hyperref}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\begin{document}

\end{document}