### 2015-12-05 comporg/review.tex
cleaning up

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{array}
\usepackage{color}
\usepackage{graphicx}
\usepackage{minted}
% \pagecolor{black}
% \color{white}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in
\begin{document}
\begin{enumerate}
\item Identify the data hazards in the following code, and fix them using $nop$ (software).
\begin{center}
\begin{minted}[linenos=true]{nasm}
sub $2, $1, $3
and $12, $2, $5
or  $13, $6, $2
add $14, $2, $2
sw  $15, 100($2)
\end{minted}
\end{center}
\indent The data hazard is present in the $\$2$ register.
The $sub$ instruction will take three cycles to write to the $\$2$ register, but the $and$ and $or$ instructions are attempting to access $\$2$ before the register has a value in it.
As a result, two $nop$ instructions should be inserted after the $sub$ instruction to offset the data access.
\begin{minted}[linenos=true]{nasm}
sub $2, $1, $3
nop
nop
and $12, $2, $5
or  $13, $6, $2
add $14, $2, $2
sw  $15, 100($2)
\end{minted}
\item Identify and fix the data hazards shown below by showing the fowarding paths (hardware).
\begin{minted}[linenos=true]{nasm}
add $3, $4, $6
sub $5, $3, $2
lw  $7, 100($5)
%add $8, $7, $2
\end{minted}
There are two data hazards in this set of instructions.
In line two, register $\$3$ is being accessed by the $sub$ instruction before it is stored by the $add$ instruction.
The other hazard is with the $\$7$ register, which is used by the $add$ instruction before a value is loaded by the $lw$ instruction.
\item Compare the performance for the single-cycle, multi-cycle, and piplined control using the following instruction mix.
\begin{center}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
& Loads & Stores & Branches & Jumps & ALU\\
\hline
Instruction Mix & 25\% & 10\% & 11\% & 2\% & 52\%\\
\hline
Clock Cycles & 5 & 4 & 3 & 3 & 4\\
\hline
\end{tabular}
\end{center}
The performance for a single cycle machine is 200ps for memory access, 100ps for ALU operations, and 50ps for register read and write.
\begin{enumerate}
\item What is the clock cycle time for a single-cycle data path?
\item What is the average CPI for the multi-cycle design?
\item What is the average CPI for the piplined design?
\item What are the average instruction times for each designs?
\end{enumerate}
\item Assume an instruction cache miss rate for a program is 2\% and a data cache miss rate is 4\%.
If the processor has a CPI of 2 without any memory stalls and the miss penalty if 100 cycles for all misses, determine how much faster a processor with a perfect cache that never misses would be.
The frequency of all loads and stores is 36\% for the SPECint2000.
\end{enumerate}
\end{document}