### 2015-12-26 pde/pdetest1.tex

\documentclass{article}
\usepackage{amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{multicol}
\usepackage{array}
\usepackage{mathrsfs}
\usepackage{minted}
\usepackage{graphics}
\usepackage{setspace}
\topmargin=-0.3in
\textheight=9.2in
\textwidth=168mm
\oddsidemargin=-0.2in
\evensidemargin=-0.2in

\newcommand{\Lap}{\mathscr{L}}
\newcommand{\Four}{\mathscr{F}}

\newcounter{mathematicapage}

\title{Partial Differential Equations}
\date{}
\begin{document}
\maketitle
\newpage
% \include{first_order}
% \newpage
% \include{classify}
% \newpage
% \include{fourier_series}
% \newpage
% \include{second_order}
% \newpage
% \include{cauchy}
% \newpage
% \include{transforms}
% \newpage
% \include{pde_formal_slns}
% \newpage
% \include{mathe_code}
% \newpage
\include{high_order}
\end{document}