### 2016-01-02 resume/education.tex

\noindent\textbf{\large Education}
\begin{enumerate}[label={\Roman*}.]
\item \textbf{Christopher Newport University $\cdot$ 2013 - Present}
\begin{itemize}
\item Pursuing a B.S. in Mathematics and a B.S. in Computer Science
\item Serving as Secretary for the Society of Physics club
\item Employed by the Center for Academic Success as a tutor
\item Earned Dean's List every semester
\item Relavent coursework includes
\begin{itemize}
\item UNIX Operating Systems, C++ Programming, Java Programming, Android Programming, Programming Language
Concepts, Data Structures, Algorithms, Computer Hardware and Design, Discreet Mathematics, Linear Algebra,
Calculus I - IV, Real Analysis, Partial and Ordinary Differential Equations, Writing in Mathematics using
\hologo{LaTeX}
\end{itemize}
\item Expected to graduate in May 2017
\end{itemize}
\end{enumerate}