### 2016-01-02 resume/work.tex

\noindent\textbf{\large Work Experience}
\begin{enumerate}[label={\Roman*}.]
\item National Aeronautics and Space Administration $\Bigcdot$ November 2015 - Present
\begin{itemize}
\item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
\item Used git extensively for version control
\item Wrote code primarily in Java
\end{itemize}
\item Freelance Application Developer $\Bigcdot$ July 2015 - Present
\begin{itemize}
\item Built a location based audio application called Audionce, written for Android and iOS.
\item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
University
\item Built websites using JavaScript, HTML5, CSS3, and PHP
\end{itemize}
\item Christopher Newport University $\Bigcdot$ September 2013 - Present
\begin{itemize}
\item IT Customer Service Representative
\begin{itemize}
\item Aided customers with various software related IT issues
\item Assisted technicians with tasks such as installing hardware
\end{itemize}
\item Tutor for the Center for Academic Success
\begin{itemize}
\item Tutored several subjects including Mathematics, Computer Engineering, and Computer Science
\item Participated in both one on one tutoring and group tutoring
\end{itemize}
\end{itemize}
\item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
\begin{itemize}
\item Lead Swim Instructor
\begin{itemize}
\item Managed a team of swim instructors that taught and oversaw various camps and lessons
\item Taught swimming lessons to children and adults of all ages
\item Promoted to manager in May 2015
\end{itemize}
\item Lifeguard
\begin{itemize}
\item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
\item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
policies of the YMCA
\end{itemize}
\end{itemize}
\end{enumerate}