### 2016-01-14 writing_in_ethics/paper_one.tex
created paper class

\documentclass{paper}

\begin{document}
\drafth{paper_one.tex}

\end{document}