### 2016-01-15 writing_in_ethics/paper_one.tex
working on paper

\documentclass{paper}

\title{Utilitarianism}
\date{}
\author{}

\addbibresource{generic_bib.bib}

\begin{document}
\drafth{}
\inlinetitle

The theory of Utilitarianism can be used to quantitatively analyze various situations that arise in day to day life; it can be applied the scenario presented in \textit{Ethics for the Information Age} that concerns Alexis' education. In order to analyze the situation and determine whether or not Alexis' actions were ethically ``good" or ``bad", the total utility change needs to be calculated. Alexis' actions can be considered ethically good if the benefit of her actions exceed the harms. Alexis' situation can be analyzed using both Rule and Act Utilitarianism.\\

Jeremy Bentham, a famous English philosopher, proposed several attributes that should be taken into account when calculating the utility of an action \cite[75]{ethics}. The first attribute, \textit{Intensity}, is concerned with the magnitude of the experience. The magnitude of Alexis' experience can be measured in the amount of happiness she experienced. She was able to achieve her goal and even positively affected others around her; presumably she was able to save her family and herself a large amount of money while also bettering her own life in the process through a college education. Both of these aspects should be taken into account as contributing positively to the magnitude of the situation.\\

The next attribute, \textit{Duration}, applies greatly to Alexis' scenario. The duration of the experience extends far beyond the time that Alexis spent in the library on the computer; in fact, the education she received will continue to affect her for the rest of her life. However, Alexis stole the information from another student at the university. The duration of that action, coupled with the amount of time she spent using the stolen information, is still greatly overshadowed by the overall happiness that Alexis experienced. It is reasonable to say that the amount of happiness gained under the \textit{Duration} criterion is far greater than the amount of happiness lost.\\

Bentham mentions the idea of certainty in his criteria for testing actions under Act Utilitarianism. There was always a chance that Alexis' end goal could have never been met, in which case the overall unhappiness produced would have far outweighed the happiness. However, Alexis did ultimately achieve her goal to be accepted into college; on top of just being accepted into college, she was also provided with a full ride, which positively affected her parents as well. As a result, the attribute of \textit{Certainty} greatly adds to the overall happiness gained from Alexis' decision. \\

The final attribute that Bentham proposes is concerned with the number of people affected by the action. Alexis' actions do in fact affect several other beings, and not all the interactions are not necessarily positive. As mentioned before, Alexis' family benefitted greatly from the situation. They were already poor; this fact may have greatly affected Alexis' ability to attend college. However, due to her actions, Alexis was able to obtain a scholarship that provided her with the funds to attend college for free.

\newpage
\printbibliography

\end{document}