### 2016-01-17 number_theory/hw/hw1.tex
homework and stuff

\documentclass{hw}

\usepackage{cancel}

\begin{document}
\makeheader{1}

\begin{enumerate}
\item What are the possible remainders when a perfect square is divided by 3 or by 6?
\begin{quote}
Let $k\in\ZZ$ such that $a = k^{2}$. By the division algorithm, we have that
\[
k = 3q_{1} + r_{1} \qquad\text{and}\qquad k = 6q_{2} + r_{2}.
\]
\underline{case $k = 3q_{1} + r_{1}$}: Squaring $k$ yields
\[
k^{2} = 9q_{1}^{2} + 6q_{1}r_{1} + r_{1}^{2}.
\]
Since $k^{2} = a$, we have that
\begin{align*}
a &= 3(3q_{1}^{2} + 2q_{1}r_{1}) + r_{1}^{2}\\
&= 3q_{0} + r_{0},\text{ where } q_{0} = (3q_{1}^{2} + 2q_{1}r_{1})\text{ and } r_{0} = r_{1}^{2}.
\end{align*}
We know by the division algorithm that $r_{0}\in\{ 0,1,2\}$. If we let $r_{0} = 0$ or $r_{0} = 1$, then
$r_{1}^{2} < 4$ and still in $\{ 0,1,2\}$. Therefore, 0 and 1 are possible remainders. If we let
$r_{0} = 2$, then $r_{1}^{2} = 4$, which is not less than 4. However, we have that
\begin{align*}
a &= 3q_{0} + 2^{2}\\
&= 3q_{0} + 4\\
&= 3q_{0} + 3 + 1\\
&= 3(q_{0} + 1) + 1,
\end{align*}
Which implies that 1 is a valid remainder. Therefore, 0 and 1 are the only valid remainders.\\

\underline{case $k = 6q_{2} + r_{2}$}:
If we square $k$, then we have
\[
k^{2} = 36q_{2}^{2} + 12q_{2}r_{2} + r_{2}^{2}.
\]
Since $a = k^{2}$, we have that
\begin{align*}
a &= 36q_{2}^{2} + 12q_{2}r_{2} + r_{2}^{2}\\
&= 6(6q_{2}^{2} + 2q_{2}r_{2}) + r_{2}^{2}\\
&= 6q_{0} + r_{0},\text{ where } q_{0} = 6q_{2}^{2} + 2q_{2}r_{2} \text{ and } r_{0} = r_{2}^{2}.
\end{align*}
We know that for $r_{0} = 0,1,\text{ and }2$, $r_{2}^{2} \leq 6$, so they are valid remainders. For
$r_{0} = 3,4,\text{ and } 5$, we have
\end{quote}
\begin{minipage}{0.3\textwidth}
\underline{$r_{0} = 3$}
\begin{align*}
a &= 6q_{0} + 9\\
&= 6q_{0} + 6 + 3\\
&= 6(q_{0} + 1) + 3
\end{align*}
$\therefore$ 3 is a valid remainder.
\end{minipage}
\begin{minipage}{0.3\textwidth}
\underline{$r_{0} = 4$}
\begin{align*}
a &= 6q_{0} + 16\\
&= 6q_{0} + 12 + 4\\
&= 6(q_{0} + 2) + 4
\end{align*}
$\therefore$ 4 is a valid remainder.
\end{minipage}
\begin{minipage}{0.3\textwidth}
\underline{$r_{0} = 5$}
\begin{align*}
a &= 6q_{0} + 25\\
&= 6q_{0} + 24 + 1\\
&= 6(q_{0} + 4) + 1
\end{align*}
$\therefore$ 1 is a valid remainder.
\end{minipage}
\begin{quote}
Therefore, the valid remainders are $\{ 0, 1, 2, 3, 4\}$.
\end{quote}
$\triangle$

\item Suppose $a,b,c,d\in\ZZ$ are such that $a|b$ and $c|d$. Prove that $ac|bd$.
\begin{quote}
Since $a|b$, we have that $b = an$ for some $n\in\ZZ$. We also have that $c|d$, so $d=cm$ for some
$m\in\ZZ$. The product $bd = (an)(cm)$, which, after rearranging, yields $bd = (nm)(ac)$. We know that
$nm\in\ZZ$, so we let $j = nm$. Then $bd = jac$, so $ac|bd$.\\
\end{quote}
$\triangle$

\item Suppose $a,b,m\in\ZZ$ and $m\neq 0$. Prove that $a|b$ if and only if $ma|mb$.
\begin{quote}
Suppose $a|b$. Then $b = na$ for some $n\in\ZZ$. Multiplying both sides of the equation yields
$mb = mna$; therefore $ma|mb$.\\
Suppose $ma|mb$, and $m\neq 0$. Then $mb = mna$ for some $n\in\ZZ$. We know that $\frac{m}{m} = 1$, so
dividing both sides by $m$ gives us $b = na$. Therefore $a|b$.
\end{quote}
$\triangle$

\item Suppose $a,b\in\ZZ$ with $b\neq 0$ and $a|b$. Prove that $|a| \leq |b|$.
\begin{quote}
Since $a|b$, we have that $b=an$ for some $n\in\ZZ$. We first need to show that $n \neq 0$.
Since $b\neq 0$, $an\neq 0$. Therefore both $a$ and $n$ must not be equal to 0. If we apply
the absolute value to $b$, we have that $|b| = |an|$. We know that $|an| = |a||n|$. Since
$n\neq 0$, $|n| \geq 1$. We also know that $|a| \geq 1$, since $a\neq 0$. Therefore
$|a||n| \geq |a|$, or $|an| \geq |a|$. Since $|b| = |an|$, we have that $|a|\leq|b|$.
\end{quote}
$\triangle$

\item Suppose that $a,b,c,d\in\ZZ$ are such that $a|b$ and $c|d$. Does $a+c$ necessarily
divide $b+d$?
\begin{quote}
Assume $a+c$ divides $b+d$. Let $a=1$ and $c=-1$. Both 1 and -1 divide all elements of $\ZZ$, but
$1 + -1 = 0$, and 0 does not divide anything. Therefore by counterexample it does not.
\end{quote}
$\triangle$

\item Suppose that $b\in\ZZ$.
\begin{enumerate}
\item If $4|b^{2}$, prove that $2|b$.
\begin{quote}
Assume that $4|b^{2}$ but $2\cancel{|}b$. Then $b = 2n + r$ for some $n$ and $r$ in $\ZZ$. We know
that $r\neq 0$, and that $0 < r < 2$. Therefore $r = 1$. If we square both sides, we get
\begin{align*}
b^{2} &= (2n+1)^{2}\\
&= 4n^{2}+4n+1\\
&= 4(n^{2} + n) + 1.
\end{align*}
Therefore $4\cancel{|}b$, since it has a remainder of 1.
\end{quote}
\item If $9|b^{2}$, prove that $3|b$.
\begin{quote}
Assume that $9|b^{2}$, but $3\cancel{|}b$. Then $b=3n+r$ where $n$ and $r$ and in $\ZZ$, and
$0 < r < 3$. If we let $r=1$, then we have
\begin{align*}
b &= 3n + 1\\
b^{2} &= (3n+1)^{2}\\
&= 9n^{2} + 6n + 1\\
&= 3(3n^{2} + 2n) + 1.
\end{align*}
Therefore $3\cancel{|}b^{2}$, and since 3 is a factor of 9, $9\cancel{|}b^{2}$.
\end{quote}
$\triangle$
\end{enumerate}

\item Use the Euclidian Algorithm to calculate $gcd(1485,1745)$.
\begin{quote}
\begin{align*}
1745 &= 1485(1) + 260\\
1485 &= 260(5) + 185\\
260 &= 185(1) + 75\\
185 &= 75(2) + 35\\
75 &= 35(2) + 5\\
35 &= 5(7) + 0\\
\end{align*}
Therefore $gcd(1485,1745) = gcd(5,0) = 5$.
\end{quote}

\item Suppose that $a$ and $b$ are positive integers. Prove that $gcd(a,b) = a$ if and
only if $a|b$.
\begin{quote}
Assume that $gcd(a,b) = a$. Then $a$ is a divisor of both $a$ and $b$ by the definition of the $gcd$.
Therefore $a|b$.
\end{quote}
\end{enumerate}
\end{document}