### 2016-01-17 vector_calc/homework/hw3.tex

\documentclass{hw}
\begin{document}
\makeheader{3}

\begin{enumerate}
\item Compute $a\cdot b$, $||a||$ and $||b||$ where $\vv{a} = 4i -3j + k$ and $\vv{b}=i+j+k$.
\begin{quote}
\begin{align*}
a\cdot b &= 4 - 3 + 1\\
&= 2\\
||a|| &= \sqrt{16+9+1}\\
&= \sqrt{26}\\
||b|| &= \sqrt{1+1+1}\\
&= \sqrt{3}
\end{align*}
\end{quote}

\item Find the angle between $\vv{a} = \sqrt{3}\ i + j$ and $\vv{a} = -\sqrt{3}\ i + j$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{\left({-2\over 4}\right)}\\
&= \arccos{\left(-{1\over 2}\right)}\\
&= {2\pi \over 3}
\end{align*}

\item Find the angle between $\vv{a} = i + j$ and $\vv{b} = i + j + k$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{\left({2 \over \sqrt{6}}\right)}
\end{align*}

\item Find the angle between $\vv{a} = (1,-2,3)$ and $\vv{b} = (3,-6,-5)$.
\begin{align*}
\theta &= \arccos{\left({a\cdot b \over ||a||\ ||b||}\right)}\\
&= \arccos{(0)}\\
&= {\pi \over 2}
\end{align*}

\item Calculate $proj_{a}b$ where $\vv{a} = i + j$ and $\vv{b} = 2i + 3j - k$.
\begin{align*}
proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
&= {4\over 2}(i + j)\\
&= 2i + 2j
\end{align*}

\item Calculate $proj_{a}b$ where $\vv{a} = i + j + 2k$ and $\vv{b} = 2i - 4j + k$.
\begin{align*}
proj_{a}b &= {a\cdot b \over ||a||^{2}}\vv{a}\\
&= {0\over 4}(i+j+2k)\\
&= 0
\end{align*}

\item Give a vector of length 3 that points in the same direction as the vector $i + j - k$.
\begin{quote}
The length is just a scalar value multiplied by the vector. Therefore we have the unit vector $i + j - k$,
with length 3 will be
\[3i + 3j - 3k.\]
\end{quote}

\item Is there ever a case where $proj_{a}b = proj_{b}a$?
\begin{quote}

\end{quote}
\end{enumerate}
\end{document}