### 2016-01-20 math_modeling/class/notes/smoothing.tex
math modeling

\documentclass{hw}

\begin{document}
\section{Exponential Smoothing}
\begin{itemize}
\item Used to predict future data trends: For example, weather forecasting
\[
S_{i} = \alpha x_{i} + (1-\alpha)S_{i-1},
\]
Where $\alpha$ and $1-\alpha$ remove the outlying data. $x_{i}$ is the actual data. This is a
recurrence relation, which can be represented by
\begin{align*}
S_{i} = \alpha(x_{i}+(1-\alpha)x_{i-1} + (1-\alpha)^{2}x_{i-2} + \dots)\\
\end{align*}
As the terms increase, the $1-\alpha$ term begins to decrease.
\end{itemize}
\end{document}