### 2016-01-25 vector_calc/homework/hw6.tex

\documentclass{hw}

\begin{document}
\makeheader{6}

\begin{enumerate}
\item Find the Cartesian coordinates of the point whose polar coordinates are given.
\begin{quote}
\begin{align*}
(\sqrt{2},{\pi\over4}) &= (\sqrt{2}\cos{\left({\pi\over4}\right)},
\sqrt{2}\sin{\left({\pi\over4}\right)})\\
&= (1,1)
\end{align*}
\end{quote}

\item Give a set of polar coordinates for the point whose Cartesian coordinates are given.
\begin{quote}
\begin{align*}
(-2,2) &= (\sqrt{(-2)^{2} + 2^{2}}, \arctan{(-1)})\\
&= (\sqrt{4},-{\pi\over4})\\
&= (2,-{\pi\over4})
\end{align*}
\end{quote}

\item Find the Cartesian coordinates of the point whose cylindrical coordinates are given.
\begin{quote}
\begin{align*}
(1,{2\pi\over3},-2) &=
(\cos{\left({2\pi\over3}\right)},\sin{\left({2\pi\over3}\right)},-2)\\
&= (-{1\over2},{\sqrt{3}\over2},-2)
\end{align*}
\end{quote}

\item Find the rectangular coordinates of the point whose spherical coordinates are given.
\begin{quote}
\begin{align*}
%rho phi theta
(1,{3\pi\over4},{2\pi\over3}) &=
(\cos{{3\pi\over4}}\sin{{2\pi\over3}},\sin{{3\pi\over4}}\sin{{2\pi\over3}},\cos{{3\pi\over4}})\\
&= ({-\sqrt{3}/2\over2},{\sqrt{3}/2\over2},{-1\over\sqrt{2}})
\end{align*}
\end{quote}

\item Find a set of cylindrical coordinates of the point whose Cartesian coordinate is given.
\begin{quote}
\begin{align*}
(-1,\sqrt{3},13) &= (\sqrt{(-1)^2 + (\sqrt{3})^2},\arctan{-{1\over\sqrt{3}}},13)\\
&= (2,-{\pi\over6},13)
\end{align*}
\end{quote}
\end{enumerate}
\end{document}