### 2016-01-31 algorithms/orders.tex
so much homework done

\documentclass{hw}

\begin{document}
\makeheader{2}

\begin{enumerate}
\item Order these functions from slowest to fastest order of growth, indicating which are the same order.
Logarithms are all base 2.
\begin{center}
\begin{tabular}{l l}
$5n^3-n^2+n$ & $O(n^3)$\\
6 & $O(1)$\\
$2^{n-1}$ & $O(2^{n})$\\
$log(log(n))$ & $O(log(n))$\\
$2^{n}$ & $O(2^n)$\\
$3^{n}$ & $O(3^n)$\\
$(log(n))^2$ & $O(n\ log(n))$\\
$n\ log(n)$ & $O(n\ log(n))$
\end{tabular}
\end{center}
\[
6 \leq log(log(n)) \leq (log(n))^{2} \leq n\ log(n) \leq 5n^3-n^2+n \leq 2^{n-1}\text{ and } 2^n \leq
3^{n}
\]

\item Find two constants $c$, $c'$ such that, for large enough $n$, we have
\[ c\ log(n) \leq log(2n^2) \leq c'\ log(n) \]
\begin{quote}
We can simplify the middle term so that the $n$ is linear for each $log$.
\begin{gather*}
c\ log(n) \leq 2log(2n) \leq c'\ log(n)\\
c\ log(n) \leq 2(log(n) + log(2)) \leq c'\ log(n)\\
c\ log(n) \leq 2 + 2log(n) \leq c'\ log(n)
\end{gather*}
If we let $c = 1$ and $c' = 4$, then for large enough $n$, the inequality will hold.
\end{quote}

\item How large must $n$ be to have $n > 13\sqrt{n}\ $?
\begin{quote}
We can solve the inequality for $n$.
\begin{gather*}
n > 13\sqrt{n}\\
n - 13\sqrt{n} > 0\\
\sqrt{n}\ (\sqrt{n} - 13) > 0
\end{gather*}
We know that $n > 0$ is not a valid, since $n = 1$ breaks the inequality. However, we also have that
$\sqrt{n} - 13 > 0$, which tells us that $n$ must be greater than $13^2$.
\end{quote}
\end{enumerate}
\end{document}