### 2016-01-31 number_theory/hw/hw3.tex

\documentclass{hw}

\begin{document}
\makeheader{3}

\begin{enumerate}
\item 
\end{enumerate}
\end{document}