### 2016-02-01 number_theory/hw/hw3.tex
working hard or something

\documentclass{hw}

\begin{document}
\makeheader{3}

\begin{enumerate}
\item Suppose that $a,b\in\ZZ$ and define the set $J={ax+by|x,y\in\ZZ}$. Prove that $J=Z$ if and only if
$gcd(a, b) = 1$.

\item We define the \textit{Fibonacci Sequence} to be the sequence of integers ${x_{0}, x_{1}, x_{2},\dots}$
satisfying the properties
\[
x_{0} = 0,\quad x_{1} = 1,\quad\text{and}\quad x_{n} = x_{n-1}+x_{n}\text{ for all } n\geq2.
\]
Prove that $gcd(x_{n},x_{n-1}) = 1$ for all $n \geq 1$. (Hint: Try using induction on $n$.)

\item Let $a, b, x$ be positive integers with $x \geq 2$ and set $d = gcd(a, b)$.
\begin{enumerate}
\item Prove that $x^{d} - 1$ divides $gcd(x^{a} - 1, x^{b} - 1)$.

\item Prove that $x^{d}-1$ is a multiple of $gcd(x^{a}-1,x^{b}-1)$ and conclude that
$x^{d}-1 = gcd(x^{a}-1,x^{b}-1)$. (Hint: We know that there exist integers $u$ and $v$ such that
$d = au + bv$. Now show that there exist integers $\alpha$ and $\beta$ such that
$x^{d}-1 = \alpha(x^{a}-1) + \beta(x^{b}-1)$.)

\end{enumerate}

\item Show that the equation $1495x + 50060y = 4$ has no solutions for $x,y\in\ZZ$.
\begin{quote}
\begin{align*}
50060 &= 1495(33) + 725\\
1495 &= 725(2) + 45\\
725 &= 45(16) + 5\\
45 &= 5(9) + 0\\
gcd(50060,1495) &= 5
\end{align*}
However, $5\cancel{|}4$, so there is no solution.
\end{quote}

\item Find all solutions to the equation $1485x + 1745y = 15$ for $x,y\in\ZZ$.

\begin{center}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
1745 &= 1485(1) + 260\\
1485 &= 260(5) + 185\\
260 &= 185(1) + 75\\
185 &= 72(2) + 35\\
75 &= 35(2) + 5\\
35 &= 7(5) + 0\\
gcd&(1745,1485) = 5
\end{align*}
\end{minipage}
\begin{minipage}[t]{0.4\textwidth}
\begin{align*}
5 &= 75 - 35(2)\\
5 &= 75 - (185 - 75(2))(2)\\
&= 75(5) - 185(2)\\
5 &= (260-185(1))(5)-185(2)\\
&= 260(5)-187(7)\\
5 &= 260(5) - (1485 -260(5))(7)\\
&= 260(40) + 1485(-7)\\
5 &= (1745 - 1485(1))(40) + 1485(-7)\\
&= 1745(40) + 1485(-47)
\end{align*}
\end{minipage}
\end{center}

$x = -47 + 349n$ and $y = 40 + 297n$ are solutions to the equation.

\item Suppose you have two small champagne glasses, one holding 8 ounces and another holding 5 ounces. Is
it possible to fill one of the glasses with exactly 1 ounce of champagne? If so, how can this be done? If
not, prove that it cannot be done.
\begin{quote}
We need to find a solution to the Diophantine Equation
\[
8x+5y=1.
\]
We know that $gcd(8,5)=1$, and $1|1$, so there is a solution. Working backwards through the Euclidian
Algorithm gives us
\begin{align*}
1 &= 3-2(1)\\
&= 3(2)-5\\
&= 8(2)+5(-3).
\end{align*}
Therefore, if we fill the first glass up twice and empty the second glass three times, we will end up with
1 ounce leftover.
\end{quote}
\end{enumerate}
\end{document}