### 2016-02-12 writing_in_ethics/paper_two.tex

\documentclass{paper}

\addbibresource{generic_bib.bib}

\title{Social Patriotism}
\begin{document}
\drafth{}
\inlinetitle
\begin{linenumbers}

Governmental power has drastically increased over the past decade. Much of the growth of power is due to acts of terrorism that have hugely impacted the nation. One of the most devastating acts of terrorism that the United States has seen is the attack on the world trade center on September 11, 2001. As a result, the government passed the USA PATRIOT act, also known as the Patriot Act. Many of the provisions in the Patriot Act grant the government the ability to avoid due process. Under the Patriot Act, the government is given excessive power to bypass the law; Kantian ethics provide an insight that can help determine whether or not use of the liberties granted to the government are ethically sound.

Section 6.6.6 of the text presents a situation in which a British news magazine, known as \textit{The Guardian}, published an article describing how the FBI indirectly requested that Verizon turn over all of its phone records. A key tenant of Kantian ethics is that others should never be used as a means to an end. The FBI's intention was never revealed, so it is hard to say whether or not they were using the phone records for personal gain or not. However, this fact alone is a breach of Kantian ethics; the FBI wanted to have access to private information, but refused to divulge the reason. Under the Categorical Imperative, moral actions should be able to be applied universally. The FBI did not release their private information, so, under the Categorical Imperative, the should not have taken others private information.

The power over communications granted to the FBI is greatly increased by the Patriot Act. Section 6.6.2 discusses the use of National Security Letters (NSLs), namely in regards to the Library Connection of Connecticut. In this scenario, the FBI used a provision of the Patriot Act to demand a user's browser history from a library computer. It follows from the argument made about the FBI requesting Verizon phone records that NSLs are a breach of Kantian ethics; the FBI prohibits individuals and companies that receive NSL's to reveal that the FBI had contacted them in the first place. This scenario differs from the Verizon scenario in that the FBI wanted the information of only one individual. The use of the NSL aside, the FBI's actions are still unethical under Kantian analysis. The agency requested that the Library Connection hand over information about another individual. In doing so, the FBI was using the Library Connection as a means to an end, which violates the Categorical Imperative. In this scenario, as well as the last, it is difficult to completely analyze the situations using Kantian analysis, since the FBI's intention is not actually know.

The third scenario present by the text involves Brandon Mayfield, an attorney from Portland, Oregon. The FBI was presented with a partial fingerprint recovered from a terrorist attack in Spain that was linked back to Mayfield. The FBI then proceeded to investigate Mayfield, going as far as entering his home and seizing private information without proper legal authorization. This scenario differs from others in that the FBI's actions were directly aimed at Mayfield. They did not use anyone as a means to an end, as they did in other scenarios. However, they still violated Kant's Categorical Imperative. Kantianism focuses on the idea that moral laws should be able to be willed into universal laws, and as such, no party involved should be able to opt-out. Following this logic, Mayfield, as well as anyone else who wanted to, would be able to take information without legal basis from the FBI. This scenario differs from the previous two in that the FBI's intention is known. The agency sought to bring a suspected terrorist into custody. The methods used to reach that goal were not ethical however, which makes their actions unethical.

It appears that under Kantian analysis, many of the actions taken by the FBI are unethical. 
\end{linenumbers}
\newpage
\printbibliography
\end{document}