### 2016-02-15 writing_in_ethics/presentation.tex
algorithm work

\documentclass{beamer}

\usepackage{url}

\title{Microtransactions and DLC}
\author{Steven Rosendahl}
\date{}

\usetheme{metropolis}

\begin{document}
\frame{\titlepage}

\begin{frame}
\frametitle{The Issue}
\begin{itemize}
\pause
\item Are micro-transactions ethical?
\pause
\item Should companies charge money for a product they market as free?
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Examples}
\pause
\begin{enumerate}
\item The Saint's Row Franchise
\pause
\begin{itemize}
\item
\end{itemize}
\end{enumerate}
\end{frame}

\begin{frame}[allowframebreaks]
\frametitle{References}
\bibliographystyle{apalike}
\bibliography{pres_bib.bib}
\end{frame}

\end{document}