### 2016-02-20 resume/work.tex

\noindent\textbf{\large Work Experience}
\begin{enumerate}[label={\Roman*}.]

\item MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present
\begin{itemize}
\item Contracted to develop Android version of iOS application
\item Collaborated with a large team of designers, managers, and programmers
\end{itemize}

\item National Aeronautics and Space Administration $\Bigcdot$ November 2015 - Present
\begin{itemize}
\item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
\item Worked closely with a university professor to plan and create the application
\end{itemize}

\item Application Developer $\Bigcdot$ July 2015 - Present
\begin{itemize}
\item Contracted by MyNewLeaf to develop Android application
\item Worked on Augmented Reality application for NASA
\item Built a location based audio application called Audionce, written for Android and iOS.
\item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
University
\item Built websites using JavaScript, HTML5, CSS3, and PHP
\end{itemize}

\item Christopher Newport University $\Bigcdot$ September 2013 - Present
\begin{itemize}
\item IT Customer Service Representative
\begin{itemize}
\item Aided customers with various software related IT issues
\item Assisted technicians with tasks such as installing hardware
\end{itemize}
\item Tutor for the Center for Academic Success
\begin{itemize}
\item Tutored several subjects including Mathematics, Computer Engineering, and Computer Science
\item Participated in both one on one tutoring and group tutoring
\end{itemize}
\end{itemize}

% \item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
% \begin{itemize}
% \item Lead Swim Instructor
% \begin{itemize}
% \item Managed a team of swim instructors that taught and oversaw various camps and lessons
% \item Taught swimming lessons to children and adults of all ages
% \item Promoted to manager in May 2015
% \end{itemize}
% \item Lifeguard
% \begin{itemize}
% \item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
% \item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
% policies of the YMCA
% \end{itemize}
% \end{itemize}
\end{enumerate}