### 2016-02-24 writing_in_ethics/abstract.tex

\documentclass{paper}

\renewcommand{\standardhead}{%
	\noindent Group B\\
	\noindent Meeting Time: MW 1500\\
	\noindent \today
}

\renewcommand{\finalh}{%
	\begin{flushright}
		\noindent Steven Rosendahl 00885793\\
		\standardhead{}
	\end{flushright}
}

\renewcommand{\inlinetitle}{\vspace{-1.5cm}{\let\newpage\relax\maketitle}\vspace{-1.5cm}}

\addbibresource{pres_bib.bib}

\date{}
\author{}
\title{Abstract}

\begin{document}
\finalh{}
\inlinetitle{}

My presentation is on the use of Downloadable Content (DLC) in order to continue making money off of gamers.
The presentation will provide concrete examples of how companies have exploited gamers using DLC. Three
ethical views, Kantianism, Utilitarianism, and Social Contract, will be used to determine if the use of
DLC is ethical or not in today's society. The presentation will argue that it is not ethical, since game
companies not only try to market incomplete products, but also lie to consumers about having a completed
product.
\nocite{*}

\printbibliography

\end{document}