### 2016-03-08 writing_in_ethics/paper_three.tex

\documentclass{paper}
\title{Title}

\addbibresource{generic_bib.bib}

\begin{document}
\drafth{}
\inlinetitle

\begin{linenumbers}

\end{linenumbers}
\newpage

\printbibliography

\end{document}