### 2016-03-10 number_theory/hw/hw6.tex
finished algorithms

\documentclass{hw}

\begin{document}
\makeheader{6}

\begin{enumerate}
\item For each of the following linear congruences of the form $ax\equiv c\mod n$, determine whether a
solution exists. If so, find a formula for all solutions and determine how many solutions there are in
$\ZZ_{n}$.
\begin{enumerate}
\item $3x\equiv 5\mod7$
\begin{quote}
The GCD of $a$ and $n$ is 1, so there is a solution since $1\mid5$, and there is only one solution.
If we solve the Diophantine equation $3x+7y=5$, we get a general solution for x as $x=3+7n$. The solution
we want is 3, since all other $n$ give a solution equal to 3 in $\ZZ_{7}$.
\end{quote}
\item $4x\equiv 9\mod12$
\begin{quote}
This congruence has no solution since $gcd(4,12)=4$, but $4\nmid9$.
\end{quote}
\item $18x\equiv 27\mod45$
\begin{quote}
The $gcd(18,45)=9$, and $9\mid27$, so there is a solution, and in fact there are 9 solutions. If we
solve the Diophantine equation $18x_{0}+45y_{0}=27$, we get that $x_{0}=-1+5n$. The nine solutions can
be found by starting with $n=1$ to $n=9$, and are $4,9,14,19,24,29,34,39,44$.
\end{quote}
\end{enumerate}

\item Let $S$ denote the number of solution to the linear congruence $ax\equiv c\mod20$. Prove that
$S\in\{ 0,1,2,3,4,5,10,20\}$.
\end{enumerate}
\end{document}