### 2016-03-16 math_modeling/hw/chw/hw.tex
nothing like some fun C code

\documentclass{hw}

\usepackage{minted}

\begin{document}
\begin{enumerate}
\item The following equation for the temperature $T=T(t)$ represents a spherical thermocouple with
convective conditions and includes radiation exchange with its surrounding walls
\[
c_{2}{dT\over\dt}=-(T-T_{\infty}+c_{1}(T^{4}-T^{4}_{sur})).
\]
In this equation, $c_{1}=1.27575\times10^{-10}K^{-3}$ and $c_{2}=0.991667s$. If it is assumed that
$T_{\infty}=473.15$, $T_{sur}=673.15K$, and $T(0)=298.15K$, determine the time $t_{s}$ at which
$T(t_{s})=490.85K$.
\begin{quote}
We can start by expressing our derivative in a discreet form. This gives the equation
\[
c_{2}\left({T(n+1)-T(n)\over\Delta t}\right)=-(T-T_{\infty}+c_{1}(T^{4}-T^{4}_{sur})).
\]
Solving this for $T(n+1)$ gives
\[
T(n+1)=-{\Delta t\over c_{2}}(T-T_{\infty}+c_{1}(T^{4}-T^{4}_{sur})) + T.
\]
\end{quote}
\end{enumerate}
\end{document}