### 2016-03-22 resume/education.tex

\noindent\textbf{\large Education}
\begin{enumerate}[label={\Roman*}.]
\item \textbf{Christopher Newport University $\cdot$ 2013 - Present}\\
Bachelors of Science $-$ Mathematics $\cdot$ Bachelors of Science $-$ Computer Science
\begin{itemize}
\item Community Involvement and Accomplishments
\begin{itemize}
\item Employed by the Center for Academic Success as a tutor in Mathematics and Computer Science
\item Inducted into the Virginia Mu chapter of Pi Mu Epsilon
\item Served as Secretary for the Society of Physics club
\item Earned Dean's List consistently
\end{itemize}
% \newpage
\item Relevant coursework includes
\begin{itemize}
\item Mathematical Education
\begin{quote}
Calculus I - IV, Linear Algebra, Real Analysis, Partial and Ordinary Differential Equations, Writing in
Mathematics using \hologo{LaTeX}, Proofs and Discreet Mathematics, Number Theory, Independent Study in
Partial Differential Equations
\end{quote}
\item Computer Science Education
\begin{quote}
UNIX Operating Systems, C++ Programming, Java Programming, Android Programming, Programming Language
Concepts, Data Structures, Algorithms, Computer Hardware and Design, Discreet Mathematics
\end{quote}
\end{itemize}
\end{itemize}
\end{enumerate}