### 2016-03-22 resume/work.tex

\noindent\textbf{\large Work Experience}
\begin{enumerate}[label={\Roman*}.]

\item MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present
\begin{itemize}
\item Contracted to develop Android version of iOS application
\item Collaborated with a large team of designers, managers, and programmers
\item Served as lead programmer for Android version of application
\end{itemize}

\item National Aeronautics and Space Administration (NASA) $\Bigcdot$ November 2015 - Present
\begin{itemize}
\item Developed Android applications for Google Glass, Vuzix, and Recon Jet headsets
\item Worked closely with a university professor to plan and create the application
\item Participated in laser safety training
\end{itemize}

\item Application Developer $\Bigcdot$ July 2015 - Present
\begin{itemize}
\item Built a location based audio application called Audionce, written for Android and iOS
\item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
University
\item Built websites using JavaScript, HTML5, CSS3, and PHP
\end{itemize}

\item Christopher Newport University $\Bigcdot$ September 2013 - Present
\begin{itemize}
\item IT Customer Service Representative
\begin{itemize}
\item Aided customers with various software related IT issues
\item Assisted technicians with tasks such as installing hardware
\end{itemize}
\item Tutor for the Center for Academic Success
\begin{itemize}
\item Tutored several subjects including Mathematics (mostly Calculus), Computer Engineering, and
Computer Science
\item Participated in both one on one tutoring and group tutoring
\item Created various resources for other Calculus tutors to use
\end{itemize}
\end{itemize}

\newpage
\item Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015
\begin{itemize}
\item Lead Swim Instructor
\begin{itemize}
\item Managed a team of swim instructors that taught and oversaw various camps and lessons
\item Taught swimming lessons to children and adults of all ages
\item Promoted to manager in May 2015
\end{itemize}
\item Lifeguard
\begin{itemize}
\item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
\item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
policies of the YMCA
\end{itemize}
\end{itemize}
\end{enumerate}