### 2016-03-31 writing_in_ethics/paper_four.tex

\documentclass{paper}
\title{The Code of Ethics}

\addbibresource{pfour.bib}

\begin{document}
\finalh{}
\inlinetitle

\begin{linenumbers}

\end{linenumbers}
\newpage

\nocite{*}
\printbibliography

\end{document}