### 2016-09-04 padics/hw2.tex
more padic work

\documentclass{hw}

\begin{document}
\makeheader{2}

\begin{enumerate}
\item If $F$ is a finite field, prove that the only absolute value on $F$ is the trivial absolute
value.
\begin{quote}
\textit{\textbf{Proof:}} By definition, if we take any $x\in F$ such that $x\neq0$, then there is some
integer $m\in\ZZ$ such that $x^{m}=1$, meaning that $|x|=1$. If $x=0$, then $|x|=0$, so the absolute
value is trivial.
\end{quote}

\item Suppose $p$ is prime. Prove that $|p^{n}|_{p}$ tends to 0 as $n\to\infty$.
\begin{quote}
\textit{\textbf{Proof:}} We know that $v_{p}(x)>0$ for all $x\in\ZZ$, and that $v_{p}(x)$ counts the
number of factors of $p$ in $x$. If we consider $p^n$, there are $n$ factors of $p$, meaning that
$v_{p}(p^{n})=n$. Then $|p^{n}|_{p}=p^{-v_{p}(p^{n})}=p^{-n}$, so
\[
\lim_{n\to\infty}p^{-n}=0.
\]
\end{quote}

\item Suppose that $F$ is a field and $|\ |$ is a non-Archemedian absolute value on $F$. if $c\in F$
and $r$ is a positive real number, we define the \textit{open ball (or disk) centered at c of radius r}
by
\[
B(c,r)=\{x\in F:|x-c|<r\}.
\]
If $d\in B(c,r)$, prove that $B(c,r)=B(d,r)$.
\end{enumerate}
\end{document}