### 2016-09-11 padics/hw3.tex
uggggg

\documentclass{hw}

\begin{document}
    \makeheader{3}
    \begin{enumerate}
        \item Suppose $F$ is a field with a non-Archemedian absolute value $|\cdot|$ and $\IRing{F}$ is the ring of
        integers of $F$. Prove that a point $x\in\IRing{F}$ is a unit if and only if $|x|=1.$
        \begin{quote}
            Suppose $x\in\IRing{F}$ is a unit and consider $x^{-1}\in\IRing{F}$ as the inverse of $x$. Since $x\in\IRing{F}$, we know
            \begin{align*}
                |x| &\leq 1\\
                |x^{-1}||x| &\leq |x^{-1}|\leq1 \qquad\text{(since $x^{-1}\in\IRing{F}$)}\\
                |x^{-1}x| &\leq |x^{-1}|\leq1\\
                |1| &\leq |x^{-1}| \leq1\\
                1 &\leq |x^{-1}| \leq 1.
            \end{align*}
            The only choice we have for $x^{-1}$ is 1, so
            \begin{align*}
                x^{-1} &= 1\\
                xx^{-1}&= x\\
                x&=1\\
                |x|&=1.
            \end{align*}
        \end{quote}

        \item If $n$ is a non-negative integer a $p$ is prime, define $\mathcal{I}_{n}=\{x\in\IRing{p}:|x|_{p}\leq p^{-n}\}$.
        \begin{enumerate}
            \item If $x,y\in\mathcal{I}_{n}$ prove that $x+y\in\mathcal{I}_{n}$.
            \begin{quote}
                Suppose $x,y\in\mathcal{I}_{n}$, and consider $|x+y|_{p}$. Then
                \begin{align*}
                    |x+y|_{p} &\leq \mmax{|x|_{p}}{|y|_{p}}\\
                    &\leq \mmax{p^{-n}}{p^{-n}}\\
                    &= p^{-n}
                \end{align*}
                Therefore $|x+y|_{p}\leq p^{-n}$, so $x+y\in\mathcal{I}_{n}$.
            \end{quote}
        \end{enumerate}
    \end{enumerate}
\end{document}