### 2016-09-27 padics/cont_fcn.tex
updating hella

% jbrunelle@mitre.org
% cnorthern@miter.org

\documentclass{hw}

\begin{document}

    \noindent We can define a function $f:\RR\to\RR$ by
    \[
        f(x)=
        \begin{cases}
            0, \qquad x\in\QQ\\
            1, \qquad x\in\RR\setminus\QQ
        \end{cases}.
    \]
    We know this function to be dicontinuous under the usual absolute value. In $\RR$ under the usual absolute value,
    we define a function $h$ to be continuous at a point $a$ in the domain of $h$ by
    \[
        \forall\epsilon > 0,\ \exists\delta > 0 \text{ such that } 0<|x-a|<\delta\implies|h(x)-h(a)|<\epsilon.
    \]
    We can reformulate this definition in the p-adic sense by
    \[
        \forall\epsilon > 0,\ \exists\delta > 0 \text{ such that } 0<|x-a|_{p}<\delta\implies|h(x)-h(a)|_{p}<\epsilon.
    \]
    If we refomulate our original function, then we can show that under the p-adic absolute value, the function is
    continuous at every point. We will define $g:\QQ_{p}\to\QQ_{p}$ as
    \[
        g(x)=
        \begin{cases}
            0, \qquad x\in\ZZ\\
            1, \qquad x\in\QQ_{p}\setminus\ZZ
        \end{cases}
    \]
    as an analog to the funtion $f$ we defined earlier. We will show that this function is indeed continuous.

    \begin{quote}
        \textit{\textbf{Proof:}} Let $\epsilon>0$ be given. If $g$ is continuous at a point $c$, then $\exists\delta>0$
        such that $\forall\epsilon >0$, $|x-c|_{p}<\delta\implies|g(x)-g(c)|_{p}<\epsilon$. We can determine our
        $\delta$ as follows:
        \begin{quote}
            If $a\in\IRing{p}$, then
            \[
                |g(x)-g(a)|_{p}=|g(x)-0|_{p}=|g(x)|_{p}=p^{-v_{p}(g(x))}\leq 1.
            \]
            If $a\in\QQ_{p}\setminus\IRing{p}$, then
            \[
                |g(x)-g(a)|_{p}=|g(x)-1|_{p} \leq \max{\{|g(x)|_{p},|-1|_{p}\}} \leq \max{\{1,1\}} = 1.
            \]
        \end{quote}
        We now let $\delta=1$. We need to consider the cases where $a\in\IRing{p}$ and $a\in\QQ\setminus\IRing{p}$. If
        $a\in\IRing{p}$ then we know $|g(x)-g(a)|_{p}\leq 1$. We need to show that there is nothing in the disk of
        radius 1 centerd 0 except for 0 and 1 (i.e. there is nothing between 0 and 1).

    \end{quote}
\end{document}