### 2016-09-27 religions_of_the_west/test_01.tex

\documentclass[8pt]{article}
\usepackage{extsizes}
\usepackage{enumitem}
\usepackage{amsmath}
\setlist{nosep}
\usepackage[margin=0.2in]{geometry}

\begin{document}
    \section*{Test 1 Review}
    \begin{enumerate}
        \item What is the Qur'an? What are the Hadiths?
            \begin{itemize}
                \item Qur'an is the holy book of Islam
                \begin{itemize}
                    \item Infallable revaltion of Muhammad
                    \item Every chapter begins by exahlting Allah
                    \item Accepts many of the prophets from the Old Testament, but is considered to be the true word of
                    God (smaller than New Testament)
                    \item Does not vereate Mary like Christian Bible, does not consider Jesus to be Messiah
                    \item Grants Mohammad special privledges (cannot be cursed, was allowed to be polygamist)
                    \item Text emphasizes wrath (Jihad, fighting is obligatory, idolitors should be killed)
                    \item Discusses final judgment
                \end{itemize}
                \item Hadiths are stories about Muhammad and how he acted in certain situations
                \begin{itemize}
                    \item Allows for clairification of certain situation not specified in Qur'an
                    \item Six collections
                \end{itemize}
            \end{itemize}

        \item What are the five pillars of Islam?
            \begin{enumerate}
                \item \textbf{Confession (Shahada)}: Must confess there is no God other than Allah
                \item \textbf{Prayer (Salat)}
                \begin{itemize}
                    \item Prayer is performed face down to the ground (submission)
                    \item Performed as a community five times a day (towards Mecca)
                \end{itemize}
                \item \textbf{Almsgiving (Zakat)}: Taking care of the poor is emphasized in the Qur'an.
                \item \textbf{Fasting (Sawm)}
                \begin{itemize}
                    \item Takes place during month of Rahmadan (when Qur'an was given to Mohammad)
                    \item Fasting during daylight hours
                \end{itemize}
                \item \textbf{Pilgrammage (Hajj)}: Journey to Mecca to circumnavigate the Ka'ba
            \end{enumerate}

        \item Differentiate the Sunnis, Shiites, Sufis, and Wahhabis
            \begin{enumerate}
                \item \textbf{Sunni}: Majority of Muslims
                \begin{itemize}
                    \item Take pride in the Muslim empire and believe that Islam can become world-dominating
                    \item Interpret scripture through reason and believe in pre-destination
                \end{itemize}
                \item \textbf{Shia}: Iran/Iraq Muslims
                \begin{itemize}
                    \item Believe Mahdi will reveal himself as messianic figure
                    \item Authority is given to clerics
                \end{itemize}
                \item \textbf{Sufi}: Mystical Muslims
                \begin{itemize}
                    \item Look for mystical meaning in the scripture
                    \item Believe Allah is one with humans (very controversial view)
                    \item Meditation on Mohammad's night journey
                    \item Pacifist; not looking for worldly Islam
                \end{itemize}
                \item \textbf{Wahhabi}: Terrorist ideology
                \begin{itemize}
                    \item Radical views of Islam
                    \item Will kill others in the name of Allah (even other Muslims)
                \end{itemize}
            \end{enumerate}

        \item Explain the war-like nature of Islam
            \begin{itemize}
                \item Allah gave Mohammad permission to fight against those who resisted Islam
                \item Mohammad was a warrior; took Mecca back after massing an army in Medina
                \item Qur'an says fighting is necessary (Jihad), but most Muslims consider it to be defensive
                \item War against infidels (Christians, Jews, etc...)
                \item Mujuhadeen (holy warriors) are promised gifts upon death
            \end{itemize}

        \item Relate Judaism and Islam
            \begin{itemize}
                \item Judaism follows from Isaac, Islam follows from Ishmael
                \item Both recognize the Old Testament (Islam sees Bible as sacred but flawed)
                \item Both tend to follow laws very strictly (Torah and Sharia)
                \item Jews did not accept Mohammad as a prophet
            \end{itemize}

        \item Who are the Black Muslims?
            \begin{itemize}
                \item Response to state of blacks in soceity
                \item Edward Byden - thought Islam could strengthen blacks place in society
                \item Timothy Drew - thought blacks had divine origins
                \item Elijah Mohammad
                \begin{itemize}
                    \item Believed he was a prophet of Allah and only he could interpret Qur'an
                    \item Blacks originated from tribe of Shabazz (ties to Arab culture)
                    \item White men were created by scientists wishing to defy Allah $\implies$ white men are evil
                    \item Blacks need to segregate themselves from whites
                \end{itemize}
            \end{itemize}

        \item Describe the process and problems of Hellenism among the Jewish people
            \begin{itemize}
                \item 330 B.C. - Alexander the Great
                \item Greek culture spread into the Isrealite lands
                \item Old Testament was translated into Greek
                \item Many Jews adopted the hellenistic ideas
                \item Some Jews saw Greek culture as too secular (adopting Greek culture meant abandoning the Torah)
            \end{itemize}

\newpage

        \item Who are the Pharisees, Sadducees, Zealots, and Essenes?
            \begin{itemize}
                \item \textbf{Pharisees}: The lower preisthood
                \begin{itemize}
                    \item Viewed Torah as an everchanging set of laws (adaptation to modern life)
                    \item Accepted oral tradition and adapted the Torah to fit it
                    \item Believed hellenistic/Romanized world created by Sadducees was doomed
                \end{itemize}
                \item \textbf{Sadducees}: The high priests
                \begin{itemize}
                    \item Civic and religious heads of society
                    \item Concerned with economic affairs - collected taxes and became very wealthy
                    \item Very conservative religious ideas (rejected angles/apocrypha and emphasized law of Moses)
                    \item Very liberal in secular matters (called Hellenizers) and compromised with many Roman demands
                \end{itemize}
                \item \textbf{Zealots}: Supporters of rebellion against Rome
                \begin{itemize}
                    \item Believed submission to Romans meant forsaking God
                    \item Fought many battles against Roman infantry
                \end{itemize}
                \item \textbf{Essenes}: Watchers of the Messiah
                \begin{itemize}
                    \item Removed themselves from society to await the Messiah
                    \item Small remote communities
                    \item Responsible for the Dead Sea Scrolls
                \end{itemize}
            \end{itemize}

        \item Differentiate Orthodox, Reform, and Conservative Judaism
            \begin{itemize}
                \item \textbf{Orthodox}: By the book Judaism
                \begin{itemize}
                    \item Believe that the written law is the only law
                    \item Very traditional and unbending to modern culture
                \end{itemize}
                \item \textbf{Reform}: Flexible tradition
                \begin{itemize}
                    \item Moral laws are eternal, written laws are not
                    \item Influenced by French Enlightenment
                    \item Did not look for return of the Messiah
                \end{itemize}
                \item \textbf{Conservative}: American Judaism
                \begin{itemize}
                    \item Tradition as a guide but not as the only choice
                    \item Middle ground between Orthodox and Reform
                \end{itemize}
            \end{itemize}

        \item Discuss the following concepts in Judaism: Jew, Synagogue, Kosher, and Sabbath
            \begin{itemize}
                \item \textbf{Jew}
                \begin{itemize}
                    \item Someone descended from the tribe of Isreal
                    \item A chosen person of God
                    \item Someone with a Jewish mother
                    \item Proselytes (converted Jews) must undergo Circumcision (males only) and Baptism
                \end{itemize}
                \item \textbf{Synagogue}
                \begin{itemize}
                    \item Jewish place of worship (\textbf{not} a temple)
                    \item Need 10 adults (orthodox - must be men) to form one
                    \item Arc containing Torah is the centerpiece (reding is very important)
                    \item Usually face towards Jerusalem
                    \item Prayer is very focused done together
                \end{itemize}
                \item \textbf{Kosher}
                \begin{itemize}
                    \item Food is sacred to Jewish people
                    \item Kosher laws are outlined in Deuteronomy/Leviticus
                    \item Keeps Jews separate from other culture
                \end{itemize}
                \item \textbf{Sabbath}
                \begin{itemize}
                    \item Holy day of worship (Saturday)
                    \item No work is to be done on Sabbath
                \end{itemize}
            \end{itemize}

        \item Relate the history of Anti-Semitism
            \begin{itemize}
                \item Begins with book of Ester - Persians attempt to anihillate Jewish people
                \item Theodosionus (Middle Ages) declares Judaism to be illegal (no conversion, Jews can't hold office)
                \item John Chrysostom - Jews are ``Christ Killers"
                \item 1096: Crusades kill many Jews, and church condemns them for usuary
                \item Jews blamed for Black Plague (poisoning wells)
                \item Many killed during Spanish inquisition
                \item Martin Luther - Just before death, writes ``On The Jews and Their Lies"
                \item Voltaire - Hated the church and the Jews (background of Church)
                \item Germany adopts influential anti-semetic ideas
                \item Hitler blames Jewish people for defeat in WWI (Jews wouldn't finance war)
                \item Holocaust of Jewish people during WWII
            \end{itemize}
    \end{enumerate}
\end{document}