### 2016-10-15 resume/education.tex
updated resume

\noindent\textbf{\large Education}
\begin{enumerate}[label={\Roman*}.]
	\item \textbf{Christopher Newport University $\Bigcdot$ August 2013 - May 2017}\\
	      Bachelors of Science $-$ Mathematics $\cdot$ Bachelors of Science $-$ Computer Science $\cdot$ 3.67 GPA
	      \begin{itemize}
	      	\item Community Involvement and Accomplishments
	      	      \begin{itemize}
	      	      	\item Employed by the Center for Academic Success as a tutor in Mathematics and Computer Science
	      	      	\item Inducted into the Virginia Mu chapter of Pi Mu Epsilon
	      	      	\item Served as Secretary for the Society of Physics club
	      	      	\item Served as Secretary for Math club
	      	      	\item Earned Dean's List since Fall 2013
	      	      \end{itemize}
	      	\item Relevant coursework includes
	      	      \begin{itemize}
	      	      	\item Mathematics Education
	      	      	      \begin{quote}
	      	      	      	\def\arraystretch{1.1}
	      	      	      	\begin{tabular}{l l l}
	      	      	      		Calculus I - IV                            & Linear Algebra         & Real Analysis                         \\
	      	      	      		Group Theory                               & Differential Equations & Technical Writing with \hologo{LaTeX} \\
	      	      	      		Proofs and Discreet Mathematics            & Number Theory          & \textit{p-adic} Analysis              \\
	      	      	      		Research in Partial Differential Equations & Mathematical Modeling  & Programming in Mathematica
	      	      	      	\end{tabular}
	      	      	      \end{quote}
	      	      	\item Computer Science Education
	      	      	      \begin{quote}
	      	      	      	\def\arraystretch{1.1}
	      	      	      	\begin{tabular}{l l l}
	      	      	      		UNIX Operating Systems & C++ Programming & Java Programming              \\
	      	      	      		Android Programming    & Software Design & Programming Language Concepts \\
	      	      	      		Data Structures        & Algorithms      & Computer Hardware and Design  \\
	      	      	      		Discreet Mathematics & Writing and Computer Ethics
	      	      	      	\end{tabular}
	      	      	      \end{quote}
	      	      	\item Programming Language and Framework Experience
	      	      	      \begin{quote}
	      	      	      	\def\arraystretch{1.1}
	      	      	      	\begin{tabular}{c  c  c}
	      	      	      		\textbf{Proficient} & \textbf{Advanced}     & \textbf{Intermediate} \\
	      	      	      		\hline
	      	      	      		Java/Android        & PHP                   & C\#/Unity             \\
	      	      	      		\LaTeX              & Javascript/Angular.js & Swift/iOS             \\
	      	      	      		Ruby                & C/C++                 & Objective-C           \\
	      	      	      		Lua                 & Python                &                       \\
	      	      	      		                    & Mathematica           &
	      	      	      	\end{tabular}
	      	      	      \end{quote}
	      	      \end{itemize}
	      \end{itemize}
\end{enumerate}