### 2016-10-15 resume/work.tex

\noindent\textbf{\large Work Experience}
\begin{enumerate}[label={\Roman*}.]

\item \textbf{MyNewLeaf, LLC $\Bigcdot$ January 2016 - Present}
\begin{itemize}
\item Contracted to develop Android version of iOS application
\item Collaborated with a large team of designers, managers, and programmers
\item Served as lead programmer for Android version of application
\item Contributed to UI design process
\end{itemize}

\item \textbf{National Aeronautics and Space Administration (NASA) $\Bigcdot$ November 2015 - Present}
\begin{itemize}
\item Developed Android applications for Google Glass, Vuzix M100, and Recon Jet headsets
\item Worked closely with a university professor to plan, create, and design the application
\item Created universal application for a wide range of Android powered devices
\item Participated in laser safety training and received laser operator certification

\end{itemize}

% \item \textbf{Application Developer $\Bigcdot$ July 2015 - Present}
% \begin{itemize}
% \item Built a location based audio application called Audionce, written for Android and iOS
% \item Created a quiz web-application called WhensIt for the Public History Center at Christopher Newport
% University
% \item Built websites using JavaScript, HTML5, CSS3, and PHP
% \end{itemize}

\item \textbf{Christopher Newport University $\Bigcdot$ September 2014 - Present}
\begin{itemize}
\item Tutor for the Center for Academic Success (Fall 2015 - Present)
\begin{itemize}
\item Tutored several subjects including Mathematics, Computer Engineering, and Computer Science
\item Participated in both one on one tutoring and group tutoring
\item Created various resources for other tutors
\item Participated in group training exercises to receive tutor certification 
\end{itemize}
\item IT Customer Service Representative (Fall 2014 - Spring 2016)
\begin{itemize}
\item Aided customers with various software related IT issues
\item Assisted technicians with tasks such as installing hardware
\end{itemize}
\end{itemize}

\newpage
\item \textbf{Kirk Family YMCA $\Bigcdot$ August 2009 - August 2015}
\begin{itemize}
\item Lead Swim Instructor
\begin{itemize}
\item Managed a team of swim instructors that taught and oversaw various camps and lessons
\item Taught swimming lessons to children and adults of all ages
\item Promoted to manager in May 2015
\end{itemize}
\item Lifeguard
\begin{itemize}
\item Attended trainings for CPR and AED certifications, as well as First Aid and lifesaving skills
\item Worked to create and maintain a safe environment for all patrons of the pool by enforcing rules and
policies of the YMCA
\end{itemize}
\end{itemize}
\end{enumerate}