### 2016-10-28 padics/hw5.tex
padics

\documentclass{hw}

\begin{document}
\makeheader{5}
\begin{enumerate}
	\item Let $f(x,y)=x^2 + y^2$. Observe that $(x,y)=(0,0)$ defines a solution to the equation $f(x,y)=0$. Do the remainder of this problem, we shall call $(0,0)$ the trivial solution to $f(x,y)=0$.
	      \begin{enumerate}
	      	\item Prove that $f(x,y)=0$ has no non-trivial solutions with $x,y\in\QQ$.
	      	\item Prove that there exists a prime $p$ such that $f(x,y)=0$ has a nontrivial solution with $x,y\in\QQ_{p}$.
	      	\item Verify that $f$ satisfies the Hasse Principle without referring to the Hasse-Minkowski Theorem.
	      \end{enumerate}

	\item Suppose that $\{ a_{n} \}_{n=0}^{\infty}$ is a sequence of points in $\ZZ_{p}$. Prove that the series
	      \[
	      	\sum_{n=0}^{\infty}a_{n}p^{n}
	      \]
	      converges in $\QQ_{p}$.
	      \begin{quote}
	      		\textit{\textbf{Proof:}} We know that a series will converge if and only if its sequence of partial sums converges. We can test for convergence by calculating
				\[
					\lim_{n\to\infty}\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p}.
				\]
				We can say
				\begin{align*}
					\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} &= \max\left\{%
						\left|a_{n+1}p^{n+1}\right|_{p},\left|a_{n}p^{n}\right|_{p}%
					\right\}\\
					&= \max\left\{%
						\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
					\right\}.\\
					\lim_{n\to\infty}\max\left\{%
						\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
					\right\} &= \max\left\{0,0\right\}\\
					&= 0.
				\end{align*}
				Therefore, the sequence of partial sums is Cauchy, and thus convergent, so the series converges.
	      \end{quote}

	\item Suppose that $k\in\NN$.
	      \begin{enumerate}
	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}\frac{1}{n^{k}}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.

	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}n^{k}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
	      \end{enumerate}
\end{enumerate}
\end{document}