### 2016-10-30 padics/hw5.tex
workign on padics

\documentclass{hw}

\begin{document}
\makeheader{5}
\begin{enumerate}
	\item Let $f(x,y)=x^2 + y^2$. Observe that $(x,y)=(0,0)$ defines a solution to the equation $f(x,y)=0$. Do the remainder of this problem, we shall call $(0,0)$ the trivial solution to $f(x,y)=0$.
	      \begin{enumerate}
	      	\item Prove that $f(x,y)=0$ has no non-trivial solutions with $x,y\in\QQ$.
	      	\item Prove that there exists a prime $p$ such that $f(x,y)=0$ has a nontrivial solution with $x,y\in\QQ_{p}$.
	      	\item Verify that $f$ satisfies the Hasse Principle without referring to the Hasse-Minkowski Theorem.
	      \end{enumerate}

	\item Suppose that $\{ a_{n} \}_{n=0}^{\infty}$ is a sequence of points in $\ZZ_{p}$. Prove that the series
	      \[
	      	\sum_{n=0}^{\infty}a_{n}p^{n}
	      \]
	      converges in $\QQ_{p}$.
	      \begin{quote}
	      	\textit{\textbf{Proof:}} We know that a series will converge if and only if its sequence of partial sums converges. We can test for convergence by calculating
	      	\[
	      		\lim_{n\to\infty}\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p}.
	      	\]
	      	We can say
	      	\begin{align*}
	      		\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} &\leq \max\left\{%
	      		\left|a_{n+1}p^{n+1}\right|_{p},\left|a_{n}p^{n}\right|_{p}%
	      		\right\}\\
	      		                                           &= \max\left\{%
	      		\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
	      		\right\}.\\
	      		\lim_{n\to\infty}\max\left\{%
	      		\frac{1}{p^{v_{p}(a_{n+1}) + n + 1}}, \frac{1}{p^{n+v_{p}(a_{n})}}%
	      		\right\}                                   & = \max\left\{0,0\right\} \\
	      		                                           & = 0.
	      	\end{align*}
			Since the $p$-adic absolute value is greater than or equal to 0, we know
			\[
			\lim_{n\to\infty}\left|a_{n+1}p^{n+1}-a_{n}p^{n}\right|_{p} = 0.
			\]
	      	Therefore, the sequence of partial sums is Cauchy, and thus convergent, so the series converges.
	      \end{quote}

	\item Suppose that $k\in\NN$.
	      \begin{enumerate}
	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}\frac{1}{n^{k}}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
	      	      \begin{quote}
	      	      	\textit{\textbf{Proof:}} If we can show that the sequence of partial sums does not converge, then the series will not converge. We want to evaluate (and eventually take the limit of)
	      	      	\[
	      	      		\left|
	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
	      	      		\right|_{p}.
	      	      	\]
	      	      	We know by definition that
	      	      	\begin{align*}
	      	      		\left|
	      	      		\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
	      	      		\right|_{p} &=
						\left|
							\frac{n^{k}-(n+1)^{k}}{n^{k}(n+1)^{k}}
						\right|_{p}\\
	      	      		&= p^{-
	      	      			\left[
	      	      			v_{p}(n^{k}-(n+1)^{k}) - v_{p}(n^{k})-v_{p}((n+1)^{k})
	      	      			\right]
	      	      		}.
	      	      	\end{align*}
					Through some tedious calculation, we can solve for the valuation of each term in the exponent.
					\begin{gather*}
						v_{p}\left(n^k-(n+1)^k\right) =
						\frac{
							\log{(n^k-(n+1)^k) - \log(m_{0})}
						}{
							\log{(p)}
						}\\
						v_{p}(n^{k}) =
						\frac{
							k\log{(n)}-\log{(m_{1})}
						}{\log{(p)}}\\
						v_{p}((n+1)^k) =
						\frac{
							k\log{(n+1)}-\log{(m_{2})}
						}{
							\log{(p)}
						},
					\end{gather*}
					where $m_{0},m_{1}$, and $m_{2}$ do not divide $p$. Combining these valuations as required yields
					\[
					-\left[v_{p}(n^{k}-(n+1)^{k}) - v_{p}(n^{k})-v_{p}((n+1)^{k})\right]=
					\frac{
					\log{\left(
						\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
					\right)}
					}{
						\log{p}
					} + c
					\]
					where $c$ is some constant number independant of $n$. We can now evaluate the real valued limit as $n\to\infty$.
					\[
						\lim_{n\to\infty}\frac{
						\log{\left(
							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
						\right)}}{\log{p}} + c = c+\frac{1}{\log{p}}\lim_{n\to\infty}\log{\left(
							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
						\right)}.
					\]
					The function on the inside of the limit is continuous for all $k\in\NN$, so we can determine the limit as follows:
					\begin{align*}
						\lim_{n\to\infty}\log{\left(
							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
						\right)} &= L\\
						e^{
						\lim_{n\to\infty}\log{\left(
							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
						\right)}
						} &= e^{L}\\
						\lim_{n\to\infty}e^{\log{\left(
							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
						\right)}
						} &= e^{L}\\
						\lim_{n\to\infty}\left(
							\frac{n^{k}(n+1)^{k}}{n^k-(n+1)^k}
						\right) &= e^{L}.
					\end{align*}
					For the fraction inside the limit, notice that after multiplication, the highest order term in the numerator will be $n^{2k}$, while the highest order term in the denominator will be $n^{k}$, so the limit behaves like
					\[
					\lim_{n\to\infty}\frac{n^{2k}}{n^k}=\lim_{n\to\infty}n^{k}=\infty = e^{L}.
					\]
					Taking the logarithm of both sides changes nothing, nor does dividing by $\log{p}$ where $p\neq\infty$, or adding $c$. Hence, we have that
					\[
					\lim_{n\to\infty}
					\left|
					\frac{1}{(n+1)^{k}}-\frac{1}{n^{k}}
					\right|_{p} = \infty,
					\]
					so the sequence of partial sums does \textbf{not} converge, and therefore the series does not converge.
	      	      \end{quote}

	      	\item Prove that the series
	      	      \[
	      	      	\sum_{n=1}^{\infty}n^{k}
	      	      \]
	      	      does not converge in $\QQ_{p}$ for any prime $p\neq\infty$.
	      	      \begin{quote}
	      	      	\textit{\textbf{Proof:}} We want to show that
	      	      	\[
	      	      		\lim_{n\to\infty}\left|(n+1)^{k} - n^{k}\right|_{p}\neq 0.
	      	      	\]
					Using the binomial theorem, we find that
					\[
					\left|(n+1)^{k} - n^{k}\right|_{p} =
					\left|\sum_{j=0}^{k}\binom{k}{j}n^{j} - n^{k}\right|_{p} =
					\left|\sum_{j=0}^{k-1}\binom{k}{j}n^{j}\right|_{p}
					\]
	      	      \end{quote}
	      \end{enumerate}
\end{enumerate}
\end{document}