### 2016-12-04 religions_of_the_west/test_02.tex
finished rstd

\documentclass[8pt]{article}
\usepackage{extsizes}
\usepackage{enumitem}
\usepackage{amsmath}
\setlist{nosep}
\usepackage[margin=0.2in]{geometry}

\begin{document}
\section*{Test 2 Review}
\begin{enumerate}
    \item Who is Ahura Mazda? Who is Angra Mainyu?
    \begin{enumerate}
        \item Ahura Mazda
        \begin{itemize}
            \item Creator God in Zoroastrianism
            \item Sole God of the religion (Zoroaster believed in one God)
            \item Followers worship his attributes (\textit{Amesha Spentas})
            \item \textit{Amesha Spentas} are good spirits
            \item Had evil twin spirit named Angra Mainyu
            \item Truth teller; truth was very important to Zoroastrianism
        \end{itemize}
        \item Angra Mainyu
        \begin{itemize}
            \item Evil counterpart to Ahura Mazda
            \item created Devas - evil spirits that paralleled \textit{Amesha Spentas}
            \item Attempted to lead Zoroaster astray
            \item Can also create things (evil is a creative force)
            \item Lier and deceiver (similar to Satan)
        \end{itemize}
    \end{enumerate}

    \item List some similarities between Zoroastrianism, Christianity, and post-exile Judaism.
    \begin{itemize}
        \item Central God; Zoroastrianism: Ahura Mazda, Judaism/Christianity: Yahweh
        \item Morality comes from God
        \item Dualism - contest between good an evil (only some sects of Judaism)
        \item Some sects of Judaism look for Messiah; Christianity has Jesus; Zoroastrianism looked for Saoshyant
        \item Coming of God - 1000 years
        \item Similar eschatology (final judgment, raising of the dead)
        \item Devil figure (Angra Mainyu/Satan)
    \end{itemize}

    \item Discuss what the church said about Jesus Christ at Nicea and Chalcedon.
    \begin{enumerate}
        \item Nicea
        \begin{itemize}
            \item First ecumenical council of the Church
            \item Needed to establish doctrine to unite Christians
            \item Hosted by Constantine
            \item Athanasius proposed homoousios
            \begin{itemize}
                \item Nature of Father and Son was very important topic
                \item Father and son have same nature but are not the same
                \item Father and son share the same essence.
            \end{itemize}
            \item Arius: Jesus was separate from God
            \begin{itemize}
                \item Jesus was first creation of God
                \item Jesus was not divine
                \item Rejected by council
            \end{itemize}
            \item Difference between \textit{made} (Arius) and \textit{begotten} (Athanasius)
            \item Trinity idea proposed (later rejected by enlightenment Christians)
        \end{itemize}
        \item Chalcedon
        \begin{itemize}
            \item Council met to settle disputes over Jesus' nature (divine vs. human)
            \item Greek ideas (God is separate) clashing with Hebrew ideas (God is with humanity)
            \item Leo's Tome: Jesus has two full natures in one person (Orthodox position)
            \item Solution proposed by Leo (bishop of Rome)
            \begin{itemize}
                \item Decided Jesus was fully God an fully man
                \item United in one person
                \item Hypostatic union
                \item Gave Leo position of bishop of bishops
            \end{itemize}
            \item Established Mary's position in the church (Humans < Mary < God)
            \item Gave Mary the status of mother of God
        \end{itemize}
    \end{enumerate}

    % \item Describe the evolution of the Papacy.
    % \begin{itemize}
    %     \item Peter considered to be head of Church after Jesus' death
    %     \item Bishops are established throughout the known world
    %     \item Chalcedon: Bishop of Rome (Leo) gains authority in establishing Doctrine
    %     \item Gregory the Great is the first recognized Pope
    %     % \item Innocent III and Gregory VII
    %     % Innocent III and Gregory VII, nation states, loss of papal power, Vatican I and II
    %     \item Pope reigns above kings in the middle ages
    %     \item Holy Roman empire formed by Constantine
    %     \item Church of England splits from the Catholic church over disputes (power of Pope weakened)
    %     \item Other theologians begin to reject Church's teaching (Martin Luther, John Calvin, etc.) further weakening the Pope's authority
    % \end{itemize}

    % \item What do Pelagius, Augustine, Erasmus, and Calvin believe about God's sovereignty and human freedom?
    % \begin{enumerate}
    %     \item Pelagius
    %     \begin{itemize}
    %         \item Humans have free will
    %         \item Good deeds will lead you to heaven
    %         \item Possibility to serve God
    %         \item Did not believe in God's saving grace
    %         \item Reacted against Augustine's teachings
    %     \end{itemize}
    %     \item Augustine
    %     \begin{itemize}
    %         \item Humans were predestined
    %         \item Give what you command and command what you will
    %         \item God selects certain people to receive his grace
    %         \begin{itemize}
    %             \item Not all who receive grace are selected (can lose grace)
    %         \end{itemize}
    %         \item Human beings are totally depraved (only God's grace can save)
    %         \item Only those who have been given the grace of God can follow
    %         \item Works on Earth do not matter
    %     \end{itemize}
    %     \item Erasmus
    %     \begin{itemize}
    %         \item Grace of God and works lead to salvation
    %         \item God's grace is far above the good works (do good through God's grace)
    %         \item Humans have free will to follow or reject God
    %         \item Strong focus on working with God
    %         \item Followed by Catholic church
    %         \item Salvation can be lost/reacquired
    %     \end{itemize}
    %     \item Calvin
    %     \begin{itemize}
    %         \item Humans have no free will (impunes sovereignty of God)
    %         \begin{itemize}
    %             \item Double predestination
    %             \item Some are chosen for Heaven, others for Hell
    %         \end{itemize}
    %         \item Salvation comes \textbf{only} from the grace of God
    %         \item Humans are inherently tainted with sin (total depravity)
    %         \begin{itemize}
    %             \item Every aspect is tainted with iniquity
    %         \end{itemize}
    %         \item Strong emphasis on God's role in human life
    %     \end{itemize}
    % \end{enumerate}

    \item Explain the following theories of atonement: Penal Substitution, Classical Redemption, Acceptation Theory, and Moral Influence
    \begin{enumerate}
        \item Penal Substitution
        \begin{itemize}
            \item Jesus was punished for sins of humanity
            \item Christ died in place of other people
            \item Righteous and wrathful God
            \item God forgives all sins through this sacrifice
        \end{itemize}
        \item Classical Redemption
        \begin{itemize}
            \item Emphasizes resurrection over the suffering
            \item God gains victory over devil
            \item Christ conquers death in the realm of the devil through resurrection
        \end{itemize}
        \item Acceptation Theory
        \begin{itemize}
            \item Calvinists identify with this position
            \item
        \end{itemize}
        \item Moral Influence
        \begin{itemize}
            \item Jesus' purpose was to bring moral teaching to humanity
            \item Sets moral example for self-sacrificing love
            \item Most popular theory
        \end{itemize}
    \end{enumerate}

    \item Differentiate the Catholic, Lutheran, Reformed, and Anabaptist theories of the sacraments of Baptism and the Eucharist
    \begin{enumerate}
        \item Catholic
        \begin{itemize}
            \item Baptism of infants
            \item Original Sin is forgiven through baptism
            \item Eucharist is \textit{Transubstantiation}: transforming into literal body and blood
            \item Eucharist is re-sacrificing of Christ
            \item Eucharist is central focus of the mass (sacrifice)
        \end{itemize}
        \item Lutheran
        \begin{itemize}
            \item Baptism of infants
            \item Original Sin is forgiven through baptism
            \item Eucharist is \textit{Consubstantiation}: elements are the same but Christ is present in, with, and under them
        \end{itemize}
        \item Reformed
        \begin{itemize}
            \item Baptism of infants
            \item Baptism is an oath by the congregation to bring up child in name of Christ (covenant)
            \item Baptism is Symbol of the Covenant
            \item Do not believe in baptismal regeneration
            \item Eucharist is symbolic of Christ; used to remember his sacrifice
        \end{itemize}
        \item Anabaptist
        \begin{itemize}
            \item Not a sacramental church
            \item Baptism of adults (child is innocent until adulthood)
            \item Baptism is symbolic of conversion and acceptance to Christ
            \item Eucharist is symbolic of Christ; used to remember his sacrifice
        \end{itemize}
    \end{enumerate}

    \item Discuss the religious and secular heritage of this country
    \begin{itemize}
        %chapters 1 and 3 of purple book
        \item Enlightenment philosophers followed Puritan ideals
        \begin{itemize}
            \item God's will revealed in nature
            \item Will is known by all humans regardless of position in life
            \item Preservation of the truth must be governments main job
            \item Natural laws $\to$ rights (Life, liberty, possessions)
        \end{itemize}
        \item Puritans came to new world to avoid religious persecution
        \begin{itemize}
            \item Locke believed in ideas of religious tolerance
            \item Government should protect all rights of the people
        \end{itemize}
        \item Puritans had democratic structure in churches
        \begin{itemize}
            \item Congregation was self sufficient
            \item Bound by its own covenant
            \item Federal government is a covenant by the people
            \item Breaking of the covenant leads to revolution
        \end{itemize}
        \item Puritans focused on equality
        \begin{itemize}
            \item Following example of Jesus
            \item Followers of Jesus were simple, blue collar
            \item Lead to ideas of equality in America
        \end{itemize}
        \item Puritans held property
        \begin{itemize}
            \item Goal of life was not to become rich, but to better society through contributions
            \item Puritans were very hard workers
            \item Lead to capitalistic ideas (Benjamin Franklin)
        \end{itemize}
        \item Various Enlightenment philosophers liked the ideas but not the religion
        \begin{itemize}
            \item Saw Church as oppressive
            \item Wanted to move away from hierarchy
        \end{itemize}
    \end{itemize}

    \item Explain the following: Mass, The Council of Trent, Franciscans, Jesuits, Immaculate Conception, Papal Infallibility, and Vatican II
    \begin{enumerate}
        \item Mass
        \begin{itemize}
            \item Religious service for Catholics
            \item Focused mainly on Eucharist (transubstantiation)
            \item Was traditionally performed in Latin or Greek; later moved to vernacular
        \end{itemize}
        \item Council of Trent
        \begin{itemize}
            \item Condemned certain beliefs of protestants
            \item Council to make certain practices of the church into official dogma
            \begin{itemize}
                \item Transubstantiation, Baptismal regeneration
            \end{itemize}
            \item Catholics follow scripture \textit{and} tradition
            \item Latin Vulgate is the accepted Bible
            \item Catholic church has sole right of scriptural interpretation
            \item Justification rests on faith, but not faith alone
            \item Cleared up disputes between different sects of Catholicism
        \end{itemize}
        \item Franciscans
        \begin{itemize}
            \item Order of monks founded by Francis of Assisi
            \item Live very simple lives (no possessions)
            \item Administer to the poor
            \item Famed for animal husbandry
        \end{itemize}
        \item Jesuits
        \begin{itemize}
            \item Founded by Ignacius Loyola
            \item Felt that other orders were tainted
            \item Tend to follow the pope directly (loyalty to Rome)
            \item Soldiers of Christ
        \end{itemize}
        \item Immaculate Conception
        \begin{itemize}
            \item Mary was born without the taint of original sin
            \item Catholics believe this leaves Mary sinless
        \end{itemize}
        \item Papal Infallibility
        \begin{itemize}
            \item Pope is direct line to God when speaking from his chair (ex cathedra)
            \item Pope's decrees are final
            \item Does not mean the pope is sinless
        \end{itemize}
        \item Vatican II
        \begin{itemize}
            \item John 23rd - Paul in the 1960's
            \item Sought to bring the church in to the modern age
            \item Liberalized the church
            \item Promoted vernacular in services as opposed to Latin
            \item Discussed relationships with other religions
            \begin{itemize}
                \item Accepted other religions
                \item Did away with antisemitism
            \end{itemize}
            \item Less authority in the papacy
        \end{itemize}
    \end{enumerate}

    \item Who are the Methodists, Mormons, Quakers, and Fundamentalists?
    \begin{enumerate}
        \item Methodists
        \begin{itemize}
            \item Founded by John Wesley
            \item Assurance of salvation
            \item Doctrine is not as important as living a good life
            \item Theology focuses on the relationship between faith and character
            \item Focus on works that lead to salvation
        \end{itemize}
        \item Mormons
        \begin{itemize}
            \item Founded by Joseph Smith
            \begin{itemize}
                \item Says he experienced visitation from God
                \item Saw corruption of church and bible
                \item Received his revelation from angel Maroni
            \end{itemize}
            \item Church of Latter Day Saints
            \item Book of Mormon
            \item Believe their leader is a prophet of God
            \item Believe that Jesus came to America
        \end{itemize}
        \item Quakers
        \begin{itemize}
            \item Founded during puritan revolution by George Fox
            \begin{itemize}
                \item Outgrowth of puritan theology
            \end{itemize}
            \item William Penn is major figure
            \item Focus on pacifism
            \item Focus on finding God in each person
            \item Focus on equality (men and women, white and black)
            \item Live removed lives
            \item Very simple dress/lifestyle
        \end{itemize}
        \item Fundamentalists
        \begin{itemize}
            \item Reaction to modernization of the church
            \item Very conservative views
            \begin{itemize}
                \item Liberals began to reject liter aspects of the Bible
                \item Proposed 5 fundamentals
                \begin{enumerate}
                    \item Scripture is infallible
                    \item Jesus was born of a virgin
                    \item Christ's death was an atonement of sin
                    \item Jesus was raised from the dead
                    \item Jesus actually performed miracles
                \end{enumerate}
            \end{itemize}
            \item Believe that many theologians had strayed from the true teaching
            \item Sought to reform teachings by referring to literal text in the Bible
        \end{itemize}
    \end{enumerate}

    \item Explain Taoism
    \begin{itemize}
        \item Founded by Lao Tzu
        \begin{itemize}
            \item Saw the ethical followers of Confucianism as too proud
            \item Believed that ethics lead to people who believed they were better than others
            \item Believed that society had forced people to become aggressive
            \item Thought that people were naturally free
        \end{itemize}
        \item Wu We: \textit{not doing}
        \begin{itemize}
            \item Taoism focuses on letting things go
            \item Letting people have their own lives is very important
        \end{itemize}
        \item The Dao is the driving force in life
        \begin{itemize}
            \item Separate from humans
            \item Dao is balanced - Humans should try to be balanced
            \item Yin and Yang is symbol of the balance of life
        \end{itemize}
    \end{itemize}
\end{enumerate}
\end{document}